Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopies

### 2024-06-16 paper.tex

\documentclass[aip, jcp, draft, onecolumn]{revtex4-2}

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of three and four wave mixing methods which form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008, RN335}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling between different vibrational, electronic, and vibronic states. \cite{RN307, RN281, RN342, Cho2008, RN335, Ogilvie2019, RN325} 
CMDS has resolved intramolecular vibrational, vibronic, excitonic, and electronic coupling in numerous systems. \cite{Carlson1990, RN345, RN342, RN343, RN324, Czech2015, Gaynor2017, Ogilvie2019, RN325}
A particular class of CMDS methods, coherent Raman based four wave mixing spectroscopies, can provide explicit probes of vibrational and excitonic coupling in a variety of systems.

@@

While SIVE showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
As a result, investigations of SIVE stagnated; to date, there are less than a dozen studies which report on SR-SIVE. \cite{RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, McDonnell2024, Bonn2024}
Recent reports highlight non-negligible SIVE interference in DOVE experiments as well as the suitability of SIVE pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
These reports motivate deeper analysis on the properties of SIVE spectroscopy.

Fully coherent, ultrafast probes of vibronic coupling would yield insight into processes that control ultrafast electronic relaxation in molecular and biological systems. \cite{Bredenbeck2015, Arsenault2021}
Some years ago, Cho proposed a parametric, doubly resonant infrared/visible pathway to probe vibronic coupling in isotropic systems. \cite{Cho2001}
In this method, a single infrared pulse is resonant with a vibrational mode, and a two photon absorption event from a near infrared (NIR) or visible pulse is used to generate the four wave mixing signal.
These pathways are referred to herein as DR-SIVE.
Methods analogous to DR-SIVE, where two photon absorption from the infrared pulse and single photon absorption from the visible pulse is used to generate output, i.e., stimulated Raman processes, have been reported. \cite{RN301, RN120}

@@

\section{Experimental Methods}
The ultrafast laser system has been described in detail elsewhere. \cite{RN278, Kaufman2024}
Briefly, an 80 MHz Ti-Sapphire ultrafast oscillator (Spectra-Physics Tsunami) created 35 fs seed pulses, amplified by a 1 kHz regnerative amplifier (Spectra-Physics Spitfire Ace) to approximately 0.5 W.
A mask in the regenerative amplifier stretched the input pulses to roughly 1 ps.
The resultant output was used to pump four independent OPAs, two of which are relevant for this study.
OPA1 (TOPAS-800, Light Conversion), referred to as $\omega_1$, was tuned for difference frequency generation (DFG) with an external stage (NDFG \"DF2\", Light Conversion).
OPA2 (OPA-800C, Spectra Physics), referred to as $\omega_2$, was tuned in its signal arrangement. 
Grating and crystal positions in OPA1 was externally controlled using \texttt{WinTOPAS} (Light Conversion).
Grating and BBO crystal positions in OPA2 were computer controlled through communication with 50 mm DC actuators (PMC). %todo: who tf makes PMC motors?
The OPAs were independently tuned through the \texttt{Attune} package, controlled by the \texttt{yaqc} and \texttt{Bluesky} packages. \cite{RN414, RN386, SkyeOPA, KyleOPA}
The power spectra and tuning results are found in Figure X.%todo: actually graph these
Movable retroreflectors (Thorlabs) controlled by 50 mm DC actuators (PMC) are used to measure coherence dynamics.
The beams are focused into the sample in a BOXCARS geometry.\cite{RN308, Kaufman2024}
To enforce frequency dependent temporal pulse overlap, the maximum points of a third order cross correlation between $\omega_1$ and $\omega_3$ measured in CaF$_2$ were used to generate frequency dependent zero delay values.
The nonlinear output was isolated through chopping schemes and a set of spatial apertures, directed into a monochromator (Horiba Micro-HR) and homodyne detected through a photomultiplier tube (Hamamatsu H7422-20), recorded and converted to digital output using a data acquisition unit (National Instruments).
\textit{in-situ} power corrections are performed by dividing the recorded spectra by OPA output measured by home-built pyroelectric detectors.
All spectra are reported on the intensity scale. %@dan if we should change to amp lmk

@@

To motivate the application of SIVE spectroscopy, we perform an approximate calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where $\alpha_{ij}$ is the Raman polarizability, N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
As such,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is roughly 100 times stronger than vibrational SFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

@@

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the total angular momentum and $M$ is its projection onto the space fixed quantization axis).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank $k$ spherical tensor, the matrix elements in \autoref{sivebeta} can be evaluated through the Wigner-Eckart theorem.
It is important to note that $\beta_{ijk}$ does not have complete tensor index symmetry.\cite{Andrews1978}
Specifically, it has shown that the rank one spherical tensor ($T_{q_1}^{(1)}$, with $q_1 \in \{-1, 0 ,1\}$) fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)} + T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Altmann1982, Andrews1990}
%there is some fancy way to note that theres no index constraint but i dont remember the language...
As such, its decomposition into irreducible tensor form is allowed only if the values of $q_2, q_3$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$ transitions are allowed, i.e., SIVE-IR features arise only from P or R branches (assuming the ground state is $\ket{g00}$).
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

@@

Most quantitative methods in an n$^{th}$ order CMDS experiment are used to extract $\chi^{(n)}$ values to compare the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.
To our knowledge, only vibraitonal SFG and 2D-IR spectroscopy has taken advantage of orientationally averaged $\chi^{(n)}$ values to extract transition moments. \cite{Shen90, Moilanen2009, RN245}
Previous treatments did not consider using quantitative analysis in SIVE to measure $\beta_{ijk}$ values from SIVE spectra.
It is thus useful to investigate how a simple treatment of orientational averaging can extract $\beta_{ijk}$ from $\chi^{(3)}_{IJKL}$.

@@

A key result of \autoref{betasive} is that the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$ and $\frac{\partial \mu_{l}}{\partial q_n}$ can be extracted from SIVE spectra. 
Since IR and spontaneous hyper-Raman spectroscopy report on $\Im(\chi^{(1)}) \sim \abs{\mu_{l}}^2$ and $\Im(\chi^{(5)}) \sim \abs{\beta_{ijk}}^2$,\cite{RN515} respectively, the absolute signs of $\frac{\partial \mu_{l}}{\partial q_n}$ and $\frac{\partial \beta_{ijk}}{\partial q_n}$ cannot be extracted from IR or hyper-Raman spectroscopy.
These ideas can prove useful when assessing the validity of computational methods which report on the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$.
Similar ideas were first introduced by Shen to extract the sign of $\frac{\partial \mu_{l}}{\partial q_n}$ from vSFG spectra.\cite{Shen90}
SIVE and vSFG together provide a unique picture of extracting signs for the microscopic observables that drive nonlinear output.

@@

The goal of this section is to understand DR-SIVE-IR selection rules.
We begin by inspecting the selection rules of a non-parametric DR-SIVE pathway (\autoref{fig:doubsive}I-a) in the driven limit. 
The driven limit is useful as despite the sum over states approach, it avoids implications of ultrafast dephasing of the vibronic state.\cite{Cho2001} 
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments (not to be confused with the response function; $\mu$ is used for transitions between states on the ground electronic state.). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
		R^{i}_{gv,ev'} 
		R^{j}_{ev',mn} 
		R^{k}_{mn,gv} 
		R^{l}_{gv,g0} 
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}
	}
\end{equation}
where $R^{i}_{gv,ev'}$ is the i$^{th}$ element of $\mel{g,v}{\vec{R}}{e,v'}$.
For generality, we assume that the vibronic  $\ket{e,v'}$ does not share the same vibration with $\ket{g,v}$. \cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}_{ab} = (a|\vec{M}|b)$ so that, for example,
$R^{i}_{gv,ev'} = \mel{v}{M^i_{ge}}{v'}$.\cite{Albrecht1960}
\begin{widetext}
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
		\mel{v}{M^{i}_{ge}}{v'} 
		\mel{v'}{M^{j}_{em}}{n}
		\mel{n}{M^{k}_{mg} }{v}
		\mel{v}{M^{l}_{gg}}{0}
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}	}
\end{equation}
Contracting over $m,n$ in \autoref{drgamma_notaylor_1} by invoking the two photon absorption polarizability $\Lambda$ gives \cite{McClain1977, Simpson2004, Olson2018}
\begin{equation}\label{drgamma_alpha_0}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^2} \sum_{v'} \frac{
		\mel{v}{M^{i}_{ge}}{v'} 
		\mel{v'}{\Lambda^{jk}_{eg}}{v}
		\mel{v}{M^{l}_{gg}}{0}
	}{\Delta_{g0,gv}
		\Delta_{ev', g0}
	}
\end{equation}
where ${\Lambda^{jk}_{eg}} = (e|\Lambda^{jk}|g)$. 
\end{widetext}
Since $\Lambda^{jk}$ transforms like a rank two tensor, it is clear that the electronic transition must be both one and two photon allowed for non-zero output, similar to resonant second harmonic generation spectroscopy. \cite{Heinz1982, Nguyen1986, Simpson2004}
This is identical to a selection rule in resonance hyper-Raman scattering, as derived by Chung and Ziegler. \cite{Ziegler1988}
As such, the DR-SIVE methods discussed here are coherent, resonance hyper-Raman analogues. 
This makes DR-SIVE a type of resonance IR spectroscopy, as introduced by Boyle et al.\cite{RN491} 
However, in the methods described by \autoref{drgamma_alpha_0}, $|e)$ must be accessible through one and two photon transitions from $|g)$.
Conversely, the resonance-IR method demonstrated by Boyle et al. only requires one photon transitions between $|g)$ and $|e)$.
Systems which provide strict parity induced selection rules (e.g., benzene) will be useful for assessing the validity of these selection rules and comparing the method described here and that of Boyle et al. \cite{RN491}

\subsection{Line Narrowing in SIVE Spectroscopies}
An unavoidable consequence in electronic spectroscopy is broadened lineshapes due to the inhomogeneity of electronic transitions (e.g., vibronic sidebands, molecular collisions). 
Four wave mixing methods can line-narrow transitions and remove some impact of inhomogeneous broadening.\cite{Carlson1990line}
The line-narrowing properties of DR-SIVE are investigated so that the pathways which line-narrow correlated or anti-correlated modes (\textit{vide infra}) can be identified.
SR-SIVE cannot linenarrow transitions as it is singly resonant. \cite{RN352}

In the discussion that follows, we assume that a vibrational mode in the ground state, $\ket{g,v}$, is correlated to the frequency of a vibronic $\ket{e,v}$.\cite{Carlson1990}
If the frequency of $\ket{g,v}$ is written as $\omega_{gv} + \xi$, where $\xi$ is some external strain (e.g. molecular collisions), then the frequency of $\ket{e,v}$ is $\omega_{ev} + c\xi$, where c is a shift coefficient.
Similarly, we take the frequency of $\ket{e,0}$ as $\omega_{e0} \rightarrow \omega_{e0} + \xi$.
The strains shift resonance denominators, e.g., for $\omega_{ij} \rightarrow \omega_{ij} + c\xi$, as
\begin{equation}
	\Delta_{ij}(\xi) = \omega_{ij} + c\xi -\omega_k - i\Gamma_{ij} \equiv \Delta_{ij}(0) + c\xi
\end{equation}
Without loss of generality, we assume c = 1 for correlated modes and c = -1 for anti-correlated modes.
To understand the impact of inhomogeneity on spectral output, we convolute an output density matrix element $\rho_{ab}(\xi)$ with a normalized Lorentzian distribution $P(\xi)$ of half width at half maximum $\sigma$ as
\begin{equation}\label{contour}
	\rho_{ab} = \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \rho_{ab}(\xi)
\end{equation} 
While Gaussian distributions are commonly used to describe inhomogeneous broadening,\cite{druet1979line, RN307, Ouellette1982} Lorentzian distributions have successfully described inhomogeneous broadening in a variety of spectroscopies and yield easily interpreted closed-form solutions. \cite{Penner1976, Desiderio1979, Dick83_1, Carlson1990line}
For clarity, we evaluate the linenarrowing characteristics of pathway Ia (\autoref{fig:doubsive}) in detail in Appendix.
The results for all pathways in \autoref{fig:doubsive} are summarized in \autoref{T:correlated}.

>>>>

\documentclass[aip, jcp, reprint, twocolumn]{revtex4-2}

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of three and four wave mixing methods which form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008, RN335}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling between different vibrational, electronic, and vibronic states. \cite{RN307, RN281, RN342, Cho2008, RN335, Ogilvie2019, RN325} 
CMDS has resolved anharmonicities and other intramolecular couplings in numerous systems. \cite{RN345, RN342, RN343, RN324, RN329, RN120, Czech2015, Gaynor2017, Ogilvie2019, RN325}
A particular class of CMDS methods, coherent Raman based four wave mixing spectroscopies, can provide explicit probes of vibrational and excitonic coupling in a variety of systems.

@@

While SIVE showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
As a result, investigations of SIVE stagnated; to date, there are less than a dozen studies which report on SR-SIVE. \cite{RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024}
Recent reports highlight non-negligible SIVE interference in DOVE experiments as well as the suitability of SIVE pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
These reports motivate deeper analysis on the properties of SIVE spectroscopy.

Fully coherent, ultrafast probes of vibronic coupling would yield insight into processes that control ultrafast electronic relaxation in molecular and biological systems. \cite{Bredenbeck2015, Arsenault2021}
Some years ago, Cho proposed a parametric, doubly resonant infrared/visible pathway to probe vibronic coupling in isotropic systems. \cite{Cho2001}
In this method, a single infrared pulse is resonant with a vibrational mode, and a two photon absorption event from a near infrared (NIR) or visible pulse resonant with an electronic or vibronic state is used to generate the four wave mixing signal.
These pathways are referred to herein as DR-SIVE.
Methods analogous to DR-SIVE, where two photon absorption from the infrared pulse and single photon absorption from the visible pulse is used to generate output, i.e., stimulated Raman processes, have been reported. \cite{RN301, RN120}

@@

\section{Experimental Methods}
The ultrafast laser system has been described in detail elsewhere. \cite{RN278, Kaufman2024}
Briefly, an 80 MHz Ti-Sapphire ultrafast oscillator  (Spectra-Physics Tsunami) created 35 fs seed pulses, amplified by a 1 kHz regnerative amplifier (Spectra-Physics Spitfire Ace) to approximately 0.5 W.
A mask in the regenerative amplifier stretched the input pulses to roughly 1 ps.
The resultant output was used to pump four independent OPAs, two of which are relevant for this study.
OPA1 (TOPAS-800, Light Conversion), referred to as $\omega_1$, was tuned for difference frequency generation (DFG) with an external stage (NDFG ``DF2'', Light Conversion).
OPA2 (OPA-800C, Spectra Physics), referred to as $\omega_2$, was tuned in its signal arrangement. 
Grating and crystal positions in OPA1 was externally controlled using \texttt{WinTOPAS} (Light Conversion).
Grating and BBO crystal positions in OPA2 were computer controlled through communication with 50 mm DC actuators (PMC). %todo: who tf makes PMC motors?
The OPAs were independently tuned through the \texttt{Attune} package, controlled by the \texttt{yaqc} and \texttt{Bluesky} packages. \cite{RN414, RN386, SkyeOPA, KyleOPA}
The power spectra and tuning results are found in Figure X. %todo: actually graph these
Movable retroreflectors (Thorlabs) controlled by 50 mm DC actuators (PMC) are used to measure coherence dynamics.
The beams are focused into the sample in a BOXCARS geometry.\cite{RN308, Kaufman2024}
To enforce frequency dependent temporal pulse overlap, the maximum points of a third order cross correlation between $\omega_1$ and $\omega_3$ measured in CaF$_2$ were used to generate frequency dependent zero delay values.
The nonlinear output was isolated through chopping schemes and a set of spatial apertures, directed into a monochromator (Horiba Micro-HR) and homodyne detected through a photomultiplier tube (Hamamatsu H7422-20), recorded and converted to digital output using a data acquisition unit (National Instruments).
\textit{in-situ} power corrections are performed by dividing the recorded spectra by OPA output measured by home-built pyroelectric detectors.
All spectra are reported on the intensity scale. %@dan if we should change to amp lmk

@@

To motivate the application of SIVE spectroscopy, we perform an approximate calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where $\alpha_{ij}$ is the Raman polarizability, N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

@@

The rovibrational selection rules are generated through calculating \autoref{sivebeta} in a $\ket{vJM} \equiv \ket{v} \otimes {\ket{JM}}$ basis, where $\{\ket{v}\}$ is a basis of vibrational states and $\{\ket{JM}\}$ is a basis of rotational states ($J$ is the total angular momentum and $M$ is its projection onto the space fixed quantization axis).
By writing $\vec{\mu} = T_{q_1}^{(1)}$ and $\beta_{ijk} = T_{q_1}^{(1)} + T_{q_2}^{(2)} + T_{q_3}^{(3)}$, where $T^{(k)}_{q}$ is a rank $k$ spherical tensor, the matrix elements in \autoref{sivebeta} can be evaluated through the Wigner-Eckart theorem.
It is important to note that $\beta_{ijk}$ does not have complete index symmetry.\cite{Andrews1978}
Specifically, it has shown that $T_{q_1}^{(1)}$ fully contributes to $\beta_{ijk}$ but only select $T_{q_2}^{(2)}$ and  $T_{q_3}^{(3)}$ terms contribute. \cite{Andrews1978, Altmann1982, Andrews1990}
%there is some fancy way to note that theres an index constraint but i dont remember the language... i think is fine
As such, its decomposition into irreducible tensor form is allowed only if the values of $q_2, q_3$ are constrained.
By substituting the spherical tensor definitions of $\vec{\mu}$ and $\beta_{ijk}$ into \autoref{sivebeta}, it is clear that only $\Delta J = \pm 1$ transitions are allowed, i.e., SIVE-IR features arise only from P or R branches (assuming the ground state is $\ket{g00}$).
This is expected since $\vec{\mu}$ transitions demand $\Delta J = \pm 1$.
As a result, consistent with vibrational SR-SIVE, rovibrational SR-SIVE follows the same rules as common infrared rovibrational spectroscopy.

@@

Most quantitative methods in an n$^{th}$ order CMDS experiment are used to extract $\chi^{(n)}$ values to compare the relative strength of nonlinear processes in different media. \cite{Zhu87, RN351, RN345}
The recorded $\chi^{(n)}$ values provide insight into how microscopic quantities ($\vec{\mu}, \alpha_{ij}, \beta_{ijk}$) impact nonlinear output.
To our knowledge, only vibrational SFG and 2D-IR spectroscopy has taken advantage of orientationally averaged $\chi^{(n)}$ values to extract transition moments. \cite{Shen90, Moilanen2009, RN245}
Previous treatments did not consider using quantitative analysis in SIVE to measure $\beta_{ijk}$ values from SIVE spectra.
It is thus useful to investigate how a simple treatment of orientational averaging can extract $\beta_{ijk}$ from $\chi^{(3)}_{IJKL}$.

@@

A key result of \autoref{betasive} is that the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$ and $\frac{\partial \mu_{l}}{\partial q_n}$ can be extracted from SIVE spectra. 
Since IR and spontaneous hyper-Raman spectroscopy report on $\Im(\chi^{(1)}) \sim \abs{\mu_{l}}^2$ and $\Im(\chi^{(5)}) \sim \abs{\beta_{ijk}}^2$,\cite{RN515} respectively, the absolute signs of $\frac{\partial \mu_{l}}{\partial q_n}$ and $\frac{\partial \beta_{ijk}}{\partial q_n}$ cannot be extracted from IR or hyper-Raman spectroscopy.
These ideas can prove useful when assessing the validity of computational methods which report on the sign of $\frac{\partial \beta_{ijk}}{\partial q_n}$.
Similar ideas were first introduced by Shen to extract the sign of $\frac{\partial \mu_{l}}{\partial q_n}$ from vSFG spectra.\cite{Shen90}
SIVE and vSFG together provide a unique picture of extracting signs for the microscopic quantities ($\mu, \alpha, \beta$) that drive nonlinear output.

@@

The goal of this section is to understand DR-SIVE-IR selection rules.
We begin by inspecting the selection rules of a non-parametric DR-SIVE pathway (\autoref{fig:doubsive}I-a) in the driven limit. 
The driven limit is useful as despite the sum over states approach, it avoids implications of ultrafast dephasing of the vibronic state.\cite{Cho2001} 
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments (not to be confused with the response function; $\vec{\mu}$ is used for transitions between states on the ground electronic state.). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
		R^{i}_{gv,ev'} 
		R^{j}_{ev',mn} 
		R^{k}_{mn,gv} 
		R^{l}_{gv,g0} 
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}
	}
\end{equation}
where $R^{i}_{gv,ev'}$ is the i$^{th}$ element of $\mel{g,v}{\vec{R}}{e,v'}$.
For generality, we assume that the vibronic  $\ket{e,v'}$ does not share the same vibration with $\ket{g,v}$. \cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}_{ab} = (a|\vec{M}|b)$ so that, for example,
$R^{i}_{gv,ev'} = \mel{v}{M^i_{ge}}{v'}$.\cite{Albrecht1960}
\begin{widetext}
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
		\mel{v}{M^{i}_{ge}}{v'} 
		\mel{v'}{M^{j}_{em}}{n}
		\mel{n}{M^{k}_{mg} }{v}
		\mel{v}{M^{l}_{gg}}{0}
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}	}
\end{equation}
Contracting over $m,n$ in \autoref{drgamma_notaylor_1} by invoking the two photon absorption polarizability $\Lambda$ gives \cite{McClain1977, Simpson2004, Olson2018}
\begin{equation}\label{drgamma_alpha_0}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^2} \sum_{v'} \frac{
		\mel{v}{M^{i}_{ge}}{v'} 
		\mel{v'}{\Lambda^{jk}_{eg}}{v}
		\mel{v}{M^{l}_{gg}}{0}
	}{\Delta_{g0,gv}
		\Delta_{ev', g0}
	}
\end{equation}
where ${\Lambda^{jk}_{eg}} = (e|\Lambda^{jk}|g)$. 
\end{widetext}
Since $\Lambda^{jk}$ transforms like a rank two tensor, it is clear that the electronic transition must be both one and two photon allowed for non-zero output, similar to resonant second harmonic generation spectroscopy. \cite{Heinz1982, Nguyen1986, Simpson2004}
This is identical to a selection rule in resonance hyper-Raman scattering, as derived by Chung and Ziegler. \cite{Ziegler1988}
As such, the DR-SIVE methods discussed here are coherent, resonance hyper-Raman analogues. 
This makes DR-SIVE a type of resonance IR spectroscopy, as introduced by Boyle et al.\cite{RN491} 
However, in the methods described by \autoref{drgamma_alpha_0}, $|e)$ must be accessible through one and two photon transitions from $|g)$.
Conversely, the resonance-IR method demonstrated by Boyle et al. only requires one photon transitions between $|g)$ and $|e)$.
Systems which provide strict parity induced selection rules (e.g., benzene) will be useful for assessing the validity of these selection rules and comparing the method described here and that of Boyle et al. \cite{RN491}

\subsection{Line Narrowing in SIVE Spectroscopies}
An unavoidable consequence in electronic spectroscopy is broadened lineshapes due to the inhomogeneity of electronic transitions (e.g., vibronic sidebands, molecular collisions). 
Four wave mixing methods can line-narrow transitions, i.e., resolve inhomogeneous broadening.\cite{Carlson1990line}
The line-narrowing properties of DR-SIVE are investigated so that the pathways which line-narrow correlated or anti-correlated modes (\textit{vide infra}) can be identified.
SR-SIVE cannot linenarrow transitions as it is singly resonant. \cite{RN352}

In the discussion that follows, we assume that a vibrational mode in the ground state, $\ket{g,v}$, is correlated to the frequency of a vibronic $\ket{e,v}$.\cite{Carlson1990}
If the frequency of $\ket{g,v}$ is written as $\omega_{gv} + \xi$, where $\xi$ is some external strain (e.g. molecular collisions), then the frequency of $\ket{e,v}$ is $\omega_{ev} + c\xi$, where c is a shift coefficient.
Similarly, we take the frequency of $\ket{e,0}$ as $\omega_{e0} \rightarrow \omega_{e0} + \xi$.
The strains shift resonance denominators, e.g., for $\omega_{ij} \rightarrow \omega_{ij} + c\xi$, as
\begin{equation}
	\Delta_{ij}(\xi) = \omega_{ij} + c\xi -\omega_k - i\Gamma_{ij} \equiv \Delta_{ij}(0) + c\xi
\end{equation}
Without loss of generality, we assume c = 1 for correlated modes and c = -1 for anti-correlated modes.
To understand the impact of inhomogeneity on spectral output, we convolute an output density matrix element $\rho_{ab}(\xi)$ with a normalized Lorentzian distribution $P(\xi)$ of half width at half maximum $\sigma$ as
\begin{equation}\label{contour}
	\rho_{ab} = \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \rho_{ab}(\xi)
\end{equation} 
While Gaussian distributions are used to describe inhomogeneous broadening,\cite{druet1979line, RN307, Ouellette1982} Lorentzian distributions have satisfactorily described inhomogeneous broadening in a variety of nonlinear spectroscopies and yield easily interpreted closed-form solutions to \autoref{contour}. \cite{Penner1976, Desiderio1979, Dick83_1, Carlson1990line}
For clarity, we evaluate the linenarrowing characteristics of pathway Ia (\autoref{fig:doubsive}) in detail in Appendix.
The results for all pathways in \autoref{fig:doubsive} are summarized in \autoref{T:correlated}.
